Make any stupid plot you want

### 2021-05-10 draft/sections/math.tex
closing out cite/spacing cleanups

One example of monoids are partial orderings on a set, such as seen in \autoref{fig:math:data:monoid}. Each hasse diagram of the set of weather states describes an ordering on the set; the arrow goes from the lesser value to the greater one. For example, $storm \leq rain$. In \autoref{fig:math:data:monoid}, the action $\bullet$ maps the elements of a set of weather states into itself by mapping them into other elements of the weather states. The action in  \autoref{fig:math:data:monoid:total}, represented as the arrows between the hasse diagrams of the weather states, maps the weather states to themselves; therefore the ordering of the weather states is identical on both sides of the action. The action $\bullet$ in \autoref{fig:math:data:monoid:partial} is a monotone map\cite{fongInvitationAppliedCategory2019}
\begin{equation*}
    if\, a \leq b \; then \; \bullet(a) \leq \bullet(b) \mid a,b \in \dfiber_{i}
\end{equation*}
where the structure the action preserves is the relative, rather than exact, ordering. Since groups are monoids with invertible operations, this definition of structure is broad enough to include the Steven's measurement scales\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale}.

@@

The base space \dbase\ provides a way to explicitly encode the continuity of the data, as described in \autoref{sec:intro:data:continuity}. This explicit topology is a concise way of distinguishing between visualizations that appear identical but assume different continuity, for example heat maps and images. The base space \dbase\ acts as an indexing space, as emphasized by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, to express how the records in \dtotal\ are connected to each other. As shown in \autoref{fig:math:data:base:types}, \dbase\ can have any number of dimensions and can be continuous or discrete. Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space\cite{aurouxMath131Introduction} such that every \(\dbasepoint \in \dbase\) has a corresponding fiber \(\dfiber_k\)\cite{QuotientSpaceTopology2020}.

@@

In \autoref{fig:math:data:base:div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping \(\pi: \dtotal\rightarrow \dbase\) is the closed interval \(\left[0,1\right]\). While the total space \dtotal\ may have components in \dfiber\ that describe any given point \(\dbasepoint \in \dbase\), such as \textit{time, latitude, longitude}, these labels are indexed into from \dbase\ the same as any other components. In contrast to the structural \textit{keys} with associated \textit{values} proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, our model treats keys \dbasepoint\ as a pure reference to topology. Decoupling the keys from their semantics allows the components identifying the keys to be altered, which provides for a coordinate agnostic representation of the continuity and facilitates encoding of data where the independent variable may not be clear. For example total rainfall is dependent on time of day and how much rain has already fallen; therefore changing the coordinate system should have no effect on how the records are connected to each other, as illustrated in \autoref{fig:intro:equivariance} where precipitation in inches and millimeters yield equivalent line plots.

@@

\subsubsection{Sheafs}
\label{sec:math:data:sheaf}
Dynamic visualizations require evaluating sections on different subspaces of \dbase; this can be achieved using a mathematical structure, called a sheaf \(\mathcal{O}\), for defining collections of objects\cite{ghristElementaryAppliedTopology2014, ghristHomologicalAlgebraData2018,urbanikBriefIntroductionSchemes} on mathematical spaces.  On the fiber bundle \dtotal, we can describe a sheaf as the collection of local sections \(\iota^{*}\dsection\)
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}[column sep=huge]
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
which are sections of \dtotal\ pulled back over local neighborhood \(U \subset \dtotal\) via the inclusion map \(\iota:\dtotal\rightarrow U\). The collation of sections enabled by sheafs is necessary for navigation techniques such as pan and zoom \cite{NekrasovskiEvaluationPanZoom2006} and dynamically updated visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

@@

As shown in \autoref{fig:math:graphic:rho:lookup}, a pixel \(p\) in the output space, drawn in yellow, is selected and mapped, via the renderer, into a region on \gtotal.  The region on \gtotal\ corresponds to a region \(\gbase_p \subset \gbase\) via the inverse mapping  \({\gsection_{xy}}^(p)\). The base space \gbase\ is an annulus to match the topology of the graphic idealized in \gtotal. The section  \(\gsection(\gbase_{p})\) then maps into the fiber \gfiber\ over \(\gbase_p\)to obtain the set of points in \gfiber, here represented as a table, that correspond to that section. The integral over the pixel components of this set of points in the fiber yields the color of the pixel. In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\ can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, or an openGL scene graph\cite{CarsonOpenGL1997} or a rendering engine such as cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}. Implementation of \gsection\ is out of scope for this proposal.

@@

\subsubsection {Visual Fiber Bundle \vtotal}
\label{sec:math:visual}
We introduce a visual bundle \vtotal\ to store the mappings of the data components into components of the graphic. These graphic components are implicit visualization library APIs; by making them explicit as components of the fiber we can define expectations of how these parameters behave. As with the data and graphic bundles, the visual bundle \((\vtotal,\,\dbase,\,\pi ,\,\vfiber)\) is defined by the projection map \(\pi\)
\begin{equation}
    \begin{tikzcd}[ampersand replacement=\&]
        \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
                          \& \dbase \arrow[u, "\vsection"', bend right]
    \end{tikzcd}
\end{equation}
where \vsection\ is the visual variable encoding, as described by Bertin \cite{bertinSemiologyGraphicsDiagrams2011a}, of the data section \dsection. The visual bundle \vtotal\ is the full design space\cite{cardStructureInformationVisualization1997} of possible parameters of a visualization type, such as a scatter plot or line plot. For example, one section \vsection\ of \vtotal\ is a tuple of visual values that specifies the visual characteristics of a part of a graphic.

@@

The Stevens measurement types\cite{stevensTheoryScalesMeasurement1946}, listed in \autoref{tab:math:artist:nu}, are specified in terms of groups, which are monoids with invertible operations\cite{remlingAlgebraMath5353}. We generalize to monoids to account for limitations in the types of data that can be described with the Stevens' scales \cite{johnsonPseudoMathematicsMentalSocial1936,thomasMathematizationNotMeasurement2014}

@@

We formally describe a glyph as \vmark\ applied to the regions \dbasepoint\ that map back to a set of path connected components $\dbasepath \subset \dbase$ as input 
\begin{equation}
\dbasepath = \{\dbasepathpoint \in \dbase \text{ exists } \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
\end{equation}
where the path\cite{ConnectedSpace2020}  $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1]. We define the glyph as the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$
\begin{equation}
  \begin{tikzcd}
      \gtotal \arrow[r, shift left] & \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] &  & \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
      \end{tikzcd}
  \label{eq:mark}
\end{equation}
such that for every glyph there is at least one corresponding region on \dbase, in keeping with the definition of glyph as any visually distinguishable element put forth by Ziemkiewicz and Kosara\cite{ziemkiewiczEmbeddingInformationVisualization2009}. The primitive point, line, and area marks\cite{bertinSemiologyGraphicsDiagrams2011a,carpendaleVisualRepresentationSemiology} are specially cased glyphs.

>>>>

One example of monoids are partial orderings on a set, such as seen in \autoref{fig:math:data:monoid}. Each hasse diagram of the set of weather states describes an ordering on the set; the arrow goes from the lesser value to the greater one. For example, $storm \leq rain$. In \autoref{fig:math:data:monoid}, the action $\bullet$ maps the elements of a set of weather states into itself by mapping them into other elements of the weather states. The action in \autoref{fig:math:data:monoid:total}, represented as the arrows between the hasse diagrams of the weather states, maps the weather states to themselves; therefore the ordering of the weather states is identical on both sides of the action. The action $\bullet$ in \autoref{fig:math:data:monoid:partial} is a monotone map \cite{fongInvitationAppliedCategory2019}
\begin{equation*}
    if\, a \leq b \; then \; \bullet(a) \leq \bullet(b) \mid a,b \in \dfiber_{i}
\end{equation*}
where the structure the action preserves is the relative, rather than exact, ordering. Since groups are monoids with invertible operations, this definition of structure is broad enough to include the Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale}.

@@

The base space \dbase\ provides a way to explicitly encode the continuity of the data, as described in \autoref{sec:intro:data:continuity}. This explicit topology is a concise way of distinguishing between visualizations that appear identical but assume different continuity, for example heat maps and images. The base space \dbase\ acts as an indexing space, as emphasized by Butler \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, to express how the records in \dtotal\ are connected to each other. As shown in \autoref{fig:math:data:base:types}, \dbase\ can have any number of dimensions and can be continuous or discrete. Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space \cite{aurouxMath131Introduction} such that every \(\dbasepoint \in \dbase\) has a corresponding fiber \(\dfiber_k\) \cite{QuotientSpaceTopology2020}.

@@

In \autoref{fig:math:data:base:div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping \(\pi: \dtotal\rightarrow \dbase\) is the closed interval \(\left[0,1\right]\). While the total space \dtotal\ may have components in \dfiber\ that describe any given point \(\dbasepoint \in \dbase\), such as \textit{time, latitude, longitude}, these labels are indexed into from \dbase\ the same as any other components. In contrast to the structural \textit{keys} with associated \textit{values} proposed by Munzner \cite{munznerVisualizationAnalysisDesign2014}, our model treats keys \dbasepoint\ as a pure reference to topology. Decoupling the keys from their semantics allows the components identifying the keys to be altered, which provides for a coordinate agnostic representation of the continuity and facilitates encoding of data where the independent variable may not be clear. For example total rainfall is dependent on time of day and how much rain has already fallen; therefore changing the coordinate system should have no effect on how the records are connected to each other, as illustrated in \autoref{fig:intro:equivariance} where precipitation in inches and millimeters yield equivalent line plots.

@@

\subsubsection{Sheafs}
\label{sec:math:data:sheaf}
Dynamic visualizations require evaluating sections on different subspaces of \dbase; this can be achieved using a mathematical structure, called a sheaf \(\mathcal{O}\), for defining collections of objects \cite{ghristElementaryAppliedTopology2014, ghristHomologicalAlgebraData2018,urbanikBriefIntroductionSchemes} on mathematical spaces.  On the fiber bundle \dtotal, we can describe a sheaf as the collection of local sections \(\iota^{*}\dsection\)
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}[column sep=huge]
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
which are sections of \dtotal\ pulled back over local neighborhood \(U \subset \dtotal\) via the inclusion map \(\iota:\dtotal\rightarrow U\). The collation of sections enabled by sheafs is necessary for navigation techniques such as pan and zoom \cite{NekrasovskiEvaluationPanZoom2006} and dynamically updated visualizations such as sliding windows \cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

@@

As shown in \autoref{fig:math:graphic:rho:lookup}, a pixel \(p\) in the output space, drawn in yellow, is selected and mapped, via the renderer, into a region on \gtotal.  The region on \gtotal\ corresponds to a region \(\gbase_p \subset \gbase\) via the inverse mapping  \({\gsection_{xy}}^(p)\). The base space \gbase\ is an annulus to match the topology of the graphic idealized in \gtotal. The section  \(\gsection(\gbase_{p})\) then maps into the fiber \gfiber\ over \(\gbase_p\)to obtain the set of points in \gfiber, here represented as a table, that correspond to that section. The integral over the pixel components of this set of points in the fiber yields the color of the pixel. In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\ can be a specification such as PDF \cite{bienz1993portable}, SVG \cite{quintScalable2003}, or an openGL scene graph \cite{CarsonOpenGL1997} or a rendering engine such as cairo \cite{CairographicsOrg} or AGG \cite{shemanarevAntiGrainGeometry}. Implementation of \gsection\ is out of scope for this proposal.

@@

\subsubsection {Visual Fiber Bundle \vtotal}
\label{sec:math:visual}
We introduce a visual bundle \vtotal\ to store the mappings of the data components into components of the graphic. These graphic components are implicit visualization library APIs; by making them explicit as components of the fiber we can define expectations of how these parameters behave. As with the data and graphic bundles, the visual bundle \((\vtotal,\,\dbase,\,\pi ,\,\vfiber)\) is defined by the projection map \(\pi\)
\begin{equation}
    \begin{tikzcd}[ampersand replacement=\&]
        \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
                          \& \dbase \arrow[u, "\vsection"', bend right]
    \end{tikzcd}
\end{equation}
where \vsection\ is the visual variable encoding, as described by Bertin \cite{bertinSemiologyGraphicsDiagrams2011a}, of the data section \dsection. The visual bundle \vtotal\ is the full design space \cite{cardStructureInformationVisualization1997} of possible parameters of a visualization type, such as a scatter plot or line plot. For example, one section \vsection\ of \vtotal\ is a tuple of visual values that specifies the visual characteristics of a part of a graphic.

@@

The Stevens measurement types \cite{stevensTheoryScalesMeasurement1946}, listed in \autoref{tab:math:artist:nu}, are specified in terms of groups, which are monoids with invertible operations\cite{remlingAlgebraMath5353}. We generalize to monoids to account for limitations in the types of data that can be described with the Stevens' scales \cite{johnsonPseudoMathematicsMentalSocial1936,thomasMathematizationNotMeasurement2014}

@@

We formally describe a glyph as \vmark\ applied to the regions \dbasepoint\ that map back to a set of path connected components $\dbasepath \subset \dbase$ as input 
\begin{equation}
\dbasepath = \{\dbasepathpoint \in \dbase \text{ exists } \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
\end{equation}
where the path \cite{ConnectedSpace2020}  $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1]. We define the glyph as the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$
\begin{equation}
  \begin{tikzcd}
      \gtotal \arrow[r, shift left] & \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] &  & \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
      \end{tikzcd}
  \label{eq:mark}
\end{equation}
such that for every glyph there is at least one corresponding region on \dbase, in keeping with the definition of glyph as any visually distinguishable element put forth by Ziemkiewicz and Kosara \cite{ziemkiewiczEmbeddingInformationVisualization2009}. The primitive point, line, and area marks \cite{bertinSemiologyGraphicsDiagrams2011a,carpendaleVisualRepresentationSemiology} are specially cased glyphs.

### 2021-05-10 draft/sections/next.tex
closing out cite/spacing cleanups

While TEAM specifies the components, the developers building libraries using TEAM components decide which compositions of components are semantically correct for the domain. For this reason, TEAM does not include data space transforms, as are incorporated into libraries like Tableau or ggplot, instead leaving choice of computations to implementors of the data object. TEAM's intentional ignorance of semantics also means it cannot evaluate whether a figurative glyph \cite{byrneAcquiredCodesMeaning2016} is a semantically correct choice, but it can enforce equivariance constraints of glyphs generated from data components enforcing equivariance of figurative glyphs \cite{byrneAcquiredCodesMeaning2016} generated from data components\cite{beckfeathers2014,byrneFigurativeFramesCritical2017}. TEAM also allows graphics to have a lower dimensional continuity than the source data when a retraction map from one continuity to the other exists. For example, TEAM components could transform 1D continuous segments into 0D discrete elements, e.g. bar charts or scatter plots. As with computations, it is the role of the domain specific library to determine which figurative glyphs and continuity downgrades are appropriate.

The prototype is deeply tied to Matplotlib's existing architecture, so it has not yet been worked through how the model generalizes to libraries such as R graphics\cite{murrellGraphicsThirdEdition2018}, VTK, and D3. TEAM has only been tested using PNG files rendered with AGG\cite{shemanarevAntiGrainGeometry}, but is expected to work with all the file types Matplotlib currently supports, including svg, pdf, and eps. We have not yet addressed how this framework interfaces with high performance rendering libraries such as openGL\cite{CarsonOpenGL1997} that implement different models of \gsection.

\subsection{Future Work}
More work is needed to formalize the composition operators, equivalence class \vartisteq, and the mathematical model of interactivity. We also need to implement artists that demonstrate that the model can underpin a minimally viable library, foremost an image\cite{haber1990visualization,hansen2011visualization}, a heat map\cite{wilkinsonHistoryClusterHeat2009,loua1873atlas}, and an inherently computational artist such as a boxplot\cite{wickham40YearsBoxplots2011}. In summary, the proposed scope of work is

\begin{table}[H]
    \centering
    \renewcommand{\arraystretch}{2}
    \begin{tabularx}{\textwidth}{|l|X|}\hline
    \textbf{work period} & \textbf{milestones \& tasks} \\ \hline
    April - July 2021  & prepare and submit \textbf{conference presentation} on new functionality enabled by model for \textit{SciPy}: \newline artists that do not inherit from existing Matplotlib artists, computational artists such as histograms, non tabular data, composite interactive artist \\ \hline
    June - Sept 2021  & prepare and submit \textbf{theory paper} on interactivity to \textit{TCVG or Eurovis 2022}: \newline fully work out and describe math of addition operators and lookups from graphic to data space, implement brush linked artist (shared base space) and artist that exploits sheafs \\ \hline
    May - Nov 2021 & prepare and submit \textbf{applications paper} on high dimensional to \textit{TCVG}: \newline math and implementation of computational artists, concurrent artists and data sources, non-trivial data bundles \\ \hline 
    Aug 2021 - Feb 2022 & prepare and submit \textbf{systems paper} on building domain specific libraries based on this model to \textit{Infoviz 2022}: \newline domain specific structured data, composite artists, inference of meta data components, mathematical notion of a visualization (labeled, multiple artists, etc)\\ \hline
    March 2022 & \textbf{dissertation} writing: \newline synthesize previous work on climate data, compile topological equivariant artist model work \\ \hline
    April 2022 & \textbf{defense} \\ \hline
    \end{tabularx}
    \caption{}
    \label{tab:code:schedule}
\end{table}
In acknowledgement that the schedule is optimistic, this work has various scales of data applications. We plan to apply this model to datasets with complex continuities, such as the trajectories of rats running around a maze and the positions of their limbs. We also potentially can look at large scale biology or climate datasets. The data applications could be further integrated with topological\cite{heineSurveyTopologybasedMethods2016} and functional \cite{ramsayFunctionalDataAnalysis2006a} data analysis methods. Since this model formalizes notions of structure preservation, it can serve as a good base for tools that assess quality metrics\cite{bertiniQualityMetricsHighdimensional2011a} or invariance \cite{kindlmannAlgebraicProcessVisualization2014} of visualizations with respect to graphical encoding choices. This specification of structure could also be used to develop a serialization structure tht could then be use to allow Matplotlib to interface with other visualization libraries such as open GL via shared serialization protocol. While this paper formulates visualization in terms of monoidal action homomorphisms between fiber bundles, the model lends itself to a categorical formulation\cite{fongInvitationAppliedCategory2019,milewskiCategoryTheoryProgrammers} that could be further explored.

>>>>

While TEAM specifies the components, the developers building libraries using TEAM components decide which compositions of components are semantically correct for the domain. For this reason, TEAM does not include data space transforms, as are incorporated into libraries like Tableau or ggplot, instead leaving choice of computations to implementors of the data object. TEAM's intentional ignorance of semantics also means it cannot evaluate whether a figurative glyph \cite{byrneAcquiredCodesMeaning2016} is a semantically correct choice, but it can enforce equivariance constraints of glyphs generated from data components enforcing equivariance of figurative glyphs \cite{byrneAcquiredCodesMeaning2016} generated from data components \cite{beckfeathers2014,byrneFigurativeFramesCritical2017}. TEAM also allows graphics to have a lower dimensional continuity than the source data when a retraction map from one continuity to the other exists. For example, TEAM components could transform 1D continuous segments into 0D discrete elements, e.g. bar charts or scatter plots. As with computations, it is the role of the domain specific library to determine which figurative glyphs and continuity downgrades are appropriate.

The prototype is deeply tied to Matplotlib's existing architecture, so it has not yet been worked through how the model generalizes to libraries such as R graphics \cite{murrellGraphicsThirdEdition2018}, VTK, and D3. TEAM has only been tested using PNG files rendered with AGG \cite{shemanarevAntiGrainGeometry}, but is expected to work with all the file types Matplotlib currently supports, including svg, pdf, and eps. We have not yet addressed how this framework interfaces with high performance rendering libraries such as openGL \cite{CarsonOpenGL1997} that implement different models of \gsection.

\subsection{Future Work}
More work is needed to formalize the composition operators, equivalence class \vartisteq, and the mathematical model of interactivity. We also need to implement artists that demonstrate that the model can underpin a minimally viable library, foremost an image \cite{haber1990visualization,hansen2011visualization}, a heat map \cite{wilkinsonHistoryClusterHeat2009,loua1873atlas}, and an inherently computational artist such as a boxplot \cite{wickham40YearsBoxplots2011}. In summary, the proposed scope of work is

\begin{table}[H]
    \centering
    \renewcommand{\arraystretch}{2}
    \begin{tabularx}{\textwidth}{|l|X|}\hline
    \textbf{work period} & \textbf{milestones \& tasks} \\ \hline
    April - July 2021  & prepare and submit \textbf{conference presentation} on new functionality enabled by model for \textit{SciPy}: \newline artists that do not inherit from existing Matplotlib artists, computational artists such as histograms, non tabular data, composite interactive artist \\ \hline
    June - Sept 2021  & prepare and submit \textbf{theory paper} on interactivity to \textit{TCVG or Eurovis 2022}: \newline fully work out and describe math of addition operators and lookups from graphic to data space, implement brush linked artist (shared base space) and artist that exploits sheafs \\ \hline
    May - Nov 2021 & prepare and submit \textbf{applications paper} on high dimensional to \textit{TCVG}: \newline math and implementation of computational artists, concurrent artists and data sources, non-trivial data bundles \\ \hline 
    Aug 2021 - Feb 2022 & prepare and submit \textbf{systems paper} on building domain specific libraries based on this model to \textit{Infoviz 2022}: \newline domain specific structured data, composite artists, inference of meta data components, mathematical notion of a visualization (labeled, multiple artists, etc)\\ \hline
    March 2022 & \textbf{dissertation} writing: \newline synthesize previous work on climate data, compile topological equivariant artist model work \\ \hline
    April 2022 & \textbf{defense} \\ \hline
    \end{tabularx}
    \caption{}
    \label{tab:code:schedule}
\end{table}
In acknowledgement that the schedule is optimistic, this work has various scales of data applications. We plan to apply this model to datasets with complex continuities, such as the trajectories of rats running around a maze and the positions of their limbs. We also potentially can look at large scale biology or climate datasets. The data applications could be further integrated with topological \cite{heineSurveyTopologybasedMethods2016} and functional \cite{ramsayFunctionalDataAnalysis2006a} data analysis methods. Since this model formalizes notions of structure preservation, it can serve as a good base for tools that assess quality metrics \cite{bertiniQualityMetricsHighdimensional2011a} or invariance \cite{kindlmannAlgebraicProcessVisualization2014} of visualizations with respect to graphical encoding choices. This specification of structure could also be used to develop a serialization structure that could then be use to allow Matplotlib to interface with other visualization libraries such as open GL via shared serialization protocol. While this paper formulates visualization in terms of monoidal action homomorphisms between fiber bundles, the model lends itself to a categorical formulation \cite{fongInvitationAppliedCategory2019,milewskiCategoryTheoryProgrammers} that could be further explored.
